Development Of Parallel Matrix Operations

### 2011-02-15 paper/paper.tex
Added more to Gaussian elimination and some minor changes.

To solve a linear system of equations of the form 
\begin{equation}
$Ax=b$, 
\end{equation}
where $A$ is an $m\times n$ matrix, $b$ is an $m\times 1$ vector, and $x$ is the $m\times 1$ solution vector, first the matrix must be inverted. For most matrices, this presents a problem, as the inverse of a matrix is generally not easy to compute and may not even exist. Gaussian elimination is the process of performing a series of elementary row operations to the matrix $A$ to reduce it to a form, known as row-echelon form, that can be easily inverted. The matrix 
\begin{equation} 
\left[ \begin{array}{ccc|c}
a_{1,1} & a_{1,2} & a_{1,3} & b_1 \\
0 & a_{2,2} & a_{2,2} & b_2 \\
0 & 0 & a_{3,3} & b_3
\end{array} \right],
\end{equation}
where $a_{i,j}$ are the elements of $A$, and $b_i$ are the elements of $b$, is an example of the system $Ax=b$ in row-echelon form. Matrices in row-echelon form have the nice characteristic that all of the leading coefficients of each row are in a column whose number is less than the column of the leading coefficients in the rows below it and greater than those in the rows above it. In the process of Gaussian elimination with partial-pivoting, this is achieved by moving the row with the greatest leading coefficient above those with lesser coefficients, and then reducing the lower rows. In the case of square matrices whose rows are linearly independent, this reduction results in an upper-triangular form. Once the matrix is in this form, the inversion process is equivalent to back substituting the solution from the bottom of the matrix all the way to the top.

@@

@@

Sequential and Parallel Implementation
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   32609        &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   53546        & 0.609        & 0.609    & xxx        \\\hline
1000    & 2       &   15812        & 2.062        & 1.031    & -0.409     \\\hline
1000    & 4       &    8101        & 4.025        & 1.006    & -0.132     \\\hline
1000    & 8       &    4703        & 6.934        & 0.991    & -0.064     \\\hline
1000    & 8       &    4210        & 7.746        & 0.968    & -0.053     \\\hline
500     & seq     &    4797        &   xxx         & xxx       & xxx        \\\hline
500     & 1       &    2046        & 2.345         & 2.345     & xxx        \\\hline
500     & 2       &    2632        & 1.823         & 0.911      & 1.573      \\\hline
500     & 4       &    1411        & 3.400         & 0.850     & 0.586      \\\hline
500     & 7       &     953        & 5.034         & 0.719     & 0.377      \\\hline
500     & 8       &     789        & 6.080         & 0.760     & 0.298      \\\hline
100     & seq     &      71        &   xxx         & xxx       & xxx        \\\hline
100     & 1       &      62        & 1.145        & 1.145    & xxx        \\\hline
100     & 2       &      42        & 1.690        & 0.845    & 0.355     \\\hline
100     & 4       &      36        & 1.972        & 0.493    & 0.441     \\\hline
100     & 7       &      35        & 2.029        & 0.290    & 0.492     \\\hline
100     & 8       &      35        & 2.029        & 0.254    & 0.502     \\\hline
\end{tabular}
\end{center}

@@

Sequential Strassen and Parallel Strassen Implementation
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   12296           &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   12105           &   1.016       & 1.016     & xxx        \\\hline
1000    & 2       &   7139            &   1.722       & 0.861     & 0.180      \\\hline
1000    & 4       &   4184            &   2.939       & 0.735     & 0.128      \\\hline
1000    & 8       &   2467            &   4.984       & 0.712     & 0.071      \\\hline
1000    & 8       &   2372            &   5.184       & 0.648     & 0.081      \\\hline
500     & seq     &   1792            &   xxx         & xxx       & xxx        \\\hline
500     & 1       &   1787            &   1.003       & 1.003     & xxx        \\\hline
500     & 2       &   1112            &   1.612       & 0.806     & 0.245      \\\hline
500     & 4       &   713             &   2.513       & 0.628     & 0.199      \\\hline
500     & 7       &   505             &   3.549       & 0.507     & 0.163      \\\hline
500     & 8       &   502             &   3.570       & 0.446     & 0.178      \\\hline
100     & seq     &   71              &   xxx         & xxx       & xxx        \\\hline
100     & 1       &   37              &   1.919       & 1.919     & xxx        \\\hline
100     & 2       &   69              &   1.029       & 0.514     & 2.730      \\\hline
100     & 4       &   70              &   1.014       & 0.254     & 2.189      \\\hline
100     & 7       &   68              &   1.044       & 0.149     & 1.977      \\\hline
100     & 8       &   68              &   1.044       & 0.131     & 1.958      \\\hline
\end{tabular}
\end{center}

>>>>

To solve a linear system of equations of the form 
\begin{equation}
$Ax=b$, 
\end{equation}
where $A$ is an $m\times n$ matrix, $b$ is an $m\times 1$ vector, and $x$ is the $m\times 1$ solution vector, first the matrix must be inverted. For most matrices, this presents a problem, as the inverse of a matrix is generally not easy to compute and may not even exist. Gaussian elimination is the process of performing a series of elementary row operations to the matrix $A$ to reduce it to a form, known as row-echelon form, that can be easily inverted. The matrix 
\begin{equation} 
\left[ \begin{array}{ccc|c}
a_{1,1} & a_{1,2} & a_{1,3} & b_1 \\
0 & a_{2,2} & a_{2,2} & b_2 \\
0 & 0 & a_{3,3} & b_3
\end{array} \right],
\end{equation}
where $a_{i,j}$ are the elements of $A$, and $b_i$ are the elements of $b$, is an example of the system $Ax=b$ in row-echelon form. Matrices in row-echelon form have the nice characteristic that all of the leading coefficients of each row are in a column whose number is less than the column of the leading coefficients in the rows below it and greater than those in the rows above it. In the process of Gaussian elimination with partial-pivoting, this is achieved by moving the row with the greatest leading coefficient above those with lesser coefficients, and then reducing the lower rows. In the case of square matrices whose rows are linearly independent, this reduction results in an upper-triangular form. Once the matrix is in this form, the inversion process is equivalent to back substituting the solution from the bottom of the matrix all the way to the top. Parallelization of this method consists of dividing up the task of reducing the rows of the matrix below the pivot point among threads.

@@

To test each of these methods, in the case of providing a matrix, a random matrix of size $n^2$ was generated by:
\begin{verbatim}
MatrixDouble A = MatrixDouble.random(1000, 1000);
\end{verbatim}
and in the case of a vector, a random vector of size $n$ was generated by:
\begin{verbatim}
VectorDouble b = VectorDouble.random(1000);
\end{verbatim}
Over 4 runs, the time required to perform the operation was averaged.

@@

\subsubsection{Sequential and Parallel Implementation}
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   32609        &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   53546        & 0.609        & 0.609    & xxx        \\\hline
1000    & 2       &   15812        & 2.062        & 1.031    & -0.409     \\\hline
1000    & 4       &    8101        & 4.025        & 1.006    & -0.132     \\\hline
1000    & 8       &    4703        & 6.934        & 0.991    & -0.064     \\\hline
1000    & 8       &    4210        & 7.746        & 0.968    & -0.053     \\\hline
500     & seq     &    4797        &   xxx         & xxx       & xxx        \\\hline
500     & 1       &    2046        & 2.345         & 2.345     & xxx        \\\hline
500     & 2       &    2632        & 1.823         & 0.911      & 1.573      \\\hline
500     & 4       &    1411        & 3.400         & 0.850     & 0.586      \\\hline
500     & 7       &     953        & 5.034         & 0.719     & 0.377      \\\hline
500     & 8       &     789        & 6.080         & 0.760     & 0.298      \\\hline
100     & seq     &      71        &   xxx         & xxx       & xxx        \\\hline
100     & 1       &      62        & 1.145        & 1.145    & xxx        \\\hline
100     & 2       &      42        & 1.690        & 0.845    & 0.355     \\\hline
100     & 4       &      36        & 1.972        & 0.493    & 0.441     \\\hline
100     & 7       &      35        & 2.029        & 0.290    & 0.492     \\\hline
100     & 8       &      35        & 2.029        & 0.254    & 0.502     \\\hline
\end{tabular}
\end{center}

@@

\subsubsection{Sequential Strassen and Parallel Strassen Implementation}
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   12296           &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   12105           &   1.016       & 1.016     & xxx        \\\hline
1000    & 2       &   7139            &   1.722       & 0.861     & 0.180      \\\hline
1000    & 4       &   4184            &   2.939       & 0.735     & 0.128      \\\hline
1000    & 8       &   2467            &   4.984       & 0.712     & 0.071      \\\hline
1000    & 8       &   2372            &   5.184       & 0.648     & 0.081      \\\hline
500     & seq     &   1792            &   xxx         & xxx       & xxx        \\\hline
500     & 1       &   1787            &   1.003       & 1.003     & xxx        \\\hline
500     & 2       &   1112            &   1.612       & 0.806     & 0.245      \\\hline
500     & 4       &   713             &   2.513       & 0.628     & 0.199      \\\hline
500     & 7       &   505             &   3.549       & 0.507     & 0.163      \\\hline
500     & 8       &   502             &   3.570       & 0.446     & 0.178      \\\hline
100     & seq     &   71              &   xxx         & xxx       & xxx        \\\hline
100     & 1       &   37              &   1.919       & 1.919     & xxx        \\\hline
100     & 2       &   69              &   1.029       & 0.514     & 2.730      \\\hline
100     & 4       &   70              &   1.014       & 0.254     & 2.189      \\\hline
100     & 7       &   68              &   1.044       & 0.149     & 1.977      \\\hline
100     & 8       &   68              &   1.044       & 0.131     & 1.958      \\\hline
\end{tabular}
\end{center}
